---
title: 'human Resources'
date: "`r Sys.Date()`"
output:
  html_document:
    fig_height: 6
    fig_width: 12
    highlight: pygments
    theme: readable
    toc: yes
---

#Intro

Hello there! In this notebook I will look at popular HR dataset and will try to solve some problems and draw some graphs.

#Prerequisites

Libraries
```{r libs, message=FALSE, warning=FALSE}
library(ggplot2)
library(gridExtra)
library(dplyr)
library(magrittr)
library(reshape2)
library(corrplot)
library(broom)
library(scales)
library(rpart.plot)
library(rpart)
library(knitr)
library(kableExtra)
```

Now load data:
```{r}
hr_data <- read.csv('../data/HR_comma_sep.csv')
```

#EDA

##Data types, missed values

```{r typing+naming}
glimpse(hr_data)
```

Hm... 'sales' column is looking to be a department, rename. Also, will format names a bit, and then will look at missed values.
```{r rename}
hr_data %<>% rename(department = sales, 
                    work_accident = Work_accident, 
                    average_monthly_hours = average_montly_hours,
                    time_spent_company = time_spend_company)
```

```{r na}
hr_data %>% 
  sapply(function(x) sum(is.na(x))) %>% 
  tidy %>%
  kable(format = 'html', col.names = c('Variable', '№ of NA')) %>%
  kable_styling(bootstrap_options = c('condensed', 'striped'), full_width = FALSE)
```

Good, no missed values. Little retyping, since 'left', 'work_accident' and 'promotion_last_5years' are factors.

```{r factorize, message = FALSE, warning= FALSE}
hr_data %<>% mutate(promotion_last_5years = factor(promotion_last_5years, levels = c(1, 0), labels = c('Yes', 'No')),
                    left = factor(left, levels = c(1, 0), labels = c('Yes', 'No')),
                    work_accident = factor(work_accident, levels = c(1, 0), labels = c('Yes', 'No')))
```


## Density plots

First of all, summary:
```{r sum}
summary(hr_data)
```

Now, lets look at density plots for continuous values and discrete plots for discrete ones.

```{r}

```


```{r densities, fig.height = 16}
sat_level <- hr_data %>%
  mutate(bin = cut(satisfaction_level, breaks = seq(0, 1, by = .1))) %>%
  group_by(bin) %>%
  count() %>%
  ungroup() %>%
  mutate(freq = n / sum(n)) %>%
  ggplot() +
    geom_bar(aes(x = bin, y = freq), stat = 'identity', fill = 'darkgreen') +
    scale_y_continuous(breaks = seq(0, 1, by = .01)) +
    labs(x = 'Satisfaction Level', y = 'Frequency') +
    theme_bw() +
    theme(panel.grid.major.y = element_line(color = 'grey', size = .1),
          panel.grid.minor.y = element_blank())

last_eval <- hr_data %>%
  mutate(bin = cut(last_evaluation, breaks = seq(0, 1, by = .1))) %>%
  group_by(bin) %>%
  count() %>%
  ungroup() %>%
  mutate(freq = n / sum(n)) %>%
  ggplot() +
    geom_bar(aes(x = bin, y = freq), stat = 'identity', fill = 'darkgreen') +
    scale_y_continuous(breaks = seq(0, 1, by = .01)) +
    labs(x = 'Last Evaluation', y = 'Frequency') +
    theme_bw() +
    theme(panel.grid.major.y = element_line(color = 'grey', size = .1),
          panel.grid.minor.y = element_blank())

avg_hours <- hr_data %>%
  mutate(bin = cut(average_monthly_hours, breaks = seq(0, 400, by = 20))) %>%
  group_by(bin) %>%
  count() %>%
  ungroup() %>%
  mutate(freq = n / sum(n)) %>%
  ggplot() +
    geom_bar(aes(x = bin, y = freq), stat = 'identity', fill = 'darkgreen') +
    scale_y_continuous(breaks = seq(0, 1, by = .01)) +
    labs(x = 'Avg Working Hours', y = 'Frequency') +
    theme_bw() +
    theme(panel.grid.major.y = element_line(color = 'grey', size = .1),
          panel.grid.minor.y = element_blank())
department <- hr_data %>%
  group_by(department) %>%
  count() %>%
  ggplot() +
    geom_bar(aes(x = reorder(department, -n), y = n), stat = 'identity', fill = 'darkgreen') +
    labs(x = 'Time spent in company, years', y = 'N') +
    scale_y_continuous(breaks = seq(0, 7000, by = 500)) +
    theme_bw() + 
    theme(panel.grid.major.y = element_line(color = 'grey', size = .1),
          panel.grid.minor.y = element_blank())

time_spent <- hr_data %>%
  ggplot() +
    geom_bar(aes(x = time_spent_company), stat = 'count', fill = 'darkgreen') +
    scale_x_continuous(breaks = seq(min(hr_data$time_spent_company), max(hr_data$time_spent_company), by = 1)) +
    labs(x = 'Time spent in company, years', y = 'N') +
    scale_y_continuous(breaks = seq(0, 7000, by = 500)) +
    theme_bw() + 
    theme(panel.grid.major.y = element_line(color = 'grey', size = .1),
          panel.grid.minor.y = element_blank())

projects <- hr_data %>%
  ggplot() +
    geom_bar(aes(x = number_project), stat = 'count', fill = 'darkgreen') +
    scale_x_continuous(breaks = seq(min(hr_data$number_project), max(hr_data$number_project), by = 1)) +
    labs(x = '№ of projects', y = 'N') +
    scale_y_continuous(breaks = seq(0, 5000, by = 500)) +
    theme_bw() + 
    theme(panel.grid.major.y = element_line(color = 'grey', size = .1),
          panel.grid.minor.y = element_blank())

layout <- rbind(c(1, 1),
                c(2, 2),
             c(3, 3),
             c(4, 4),
             c(5, 6))
grid.arrange(grobs = list(sat_level, last_eval, avg_hours, department, time_spent, projects), layout_matrix = layout)
```

Cool! Now, our factor variables.

```{r factors}
accident <- hr_data %>% count(work_accident) %>% mutate(freq = n / sum(n)) %>%
  ggplot() +
  geom_bar(aes(x = work_accident, fill = work_accident, y = freq), stat = 'identity') +
  scale_fill_manual(values = c('darkgreen', 'firebrick'), guide = FALSE) +
  scale_y_continuous(breaks = seq(0, 1, by = .05), labels = percent) +
  xlab('Had accident on work') +
  theme_bw() + 
  theme(panel.grid.major.y = element_line(size = .1, color = 'grey'),
        panel.grid.major.x = element_blank())

promo <- hr_data %>% count(promotion_last_5years) %>% mutate(freq = n / sum(n)) %>%
  ggplot() +
  geom_bar(aes(x = promotion_last_5years, fill = promotion_last_5years, y = freq), stat = 'identity') +
  scale_fill_manual(values = c('darkgreen', 'firebrick'), guide = FALSE) +
  scale_y_continuous(breaks = seq(0, 1, by = .05), labels = percent) +
  xlab('Been promoted in last 5 years') +
  theme_bw() + 
  theme(panel.grid.major.y = element_line(size = .1, color = 'grey'),
        panel.grid.major.x = element_blank())

left <- hr_data %>% count(left) %>% mutate(freq = n / sum(n)) %>%
  ggplot() +
  geom_bar(aes(x = left, fill = left, y = freq), stat = 'identity') +
  scale_fill_manual(values = c('darkgreen', 'firebrick'), guide = FALSE) +
  scale_y_continuous(breaks = seq(0, 1, by = .05), labels = percent) +
  xlab('Left job') +
  theme_bw() + 
  theme(panel.grid.major.y = element_line(size = .1, color = 'grey'),
        panel.grid.major.x = element_blank())
  
layout <- rbind(c(1, 2, 3))
grid.arrange(grobs = list(accident, promo, left), layout_matrix = layout)
```

So, very few been promoted, about quarter of people left their job.

Ok, let's look at correlogram.
```{r corrplot}
hr_data %>% mutate(work_accident = as.numeric(work_accident),
                   left = as.numeric(left),
                   promotion_last_5years = as.numeric(promotion_last_5years),
                   department = as.numeric(department),
                   salary = as.numeric(salary)) %>% cor %>% corrplot(method = c('square'))
```

#Some sort of practical analysis

So, why people leave? Let's figure out. In lazy way, we'll make decision tree model, look at it's internals, and then will build graphs. :)

```{r trees}
model <- rpart(left~., hr_data)
rpart.plot(model)
```

So, main concern is satisfaction level. Let's see difference of distribution between left and stayed employees.


```{r sat_left, fig.width=12}
hr_data %>% 
  select(left, satisfaction_level) %>%
  mutate(bin = cut(satisfaction_level, breaks = seq(0, 1, by = .05), labels = FALSE) / 20) %>%
  group_by(bin, left) %>%
  count() %>%
  ungroup() %>%
  group_by(bin) %>%
  mutate(freq = n / sum(n)) %>%
  ggplot() +
    geom_bar(aes(x = bin, y = freq, fill = left), stat = 'identity') +
    scale_fill_manual(values = c('darkgreen', 'firebrick')) +
    geom_vline(aes(xintercept = .46), color = 'white', linetype = 'dashed', size = 1) +
    scale_color_manual(values = c('darkgreen', 'orangered')) +
    guides(fill = guide_legend(title = 'Has left job?')) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 1)) +
    labs(x = 'Satisfaction Level', y = 'Frequency') +
    theme_bw() + 
    scale_y_continuous(breaks = seq(0, 1, by = .05), labels = percent) +
    theme(panel.grid.major.y = element_line(size = .1, color = 'grey'),
          panel.grid.major.x = element_blank())
```

Dashed line is the boundary decision from a tree. Looks like we can locate 2 big groups with this decision boundary. One still left, so we go down right on the tree and check next 2 variables.

```{r heatmap}
hr_data %>% 
  filter(satisfaction_level >= .46) %>% 
  mutate(bin = ntile(last_evaluation, 10)/10) %>% 
  group_by(bin, time_spent_company) %>% 
  count(left) %>%
  mutate(freq = n / sum(n)) %>%
  filter(left == 'Yes') %>%
  ggplot(aes(bin, time_spent_company)) +
    geom_tile(aes(fill = freq)) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, 6)) +
    scale_y_continuous(breaks = seq(0, 6, by = 1)) +
    scale_x_continuous(breaks = seq(0, 1, by = .1)) +
    labs(x = 'Last evaluation', y = 'Time spent in company', fill = 'Frequency') +
    scale_fill_gradient(low = 'white', high = 'darkgreen') +
    theme_bw() +
    theme(panel.grid = element_blank())
```

Ok, another zone is localized, big possibility to leave is for men who worked for 5+ years and was highly productive by their last measures.
Judging by decision tree, we can guess that you probably leave from your job if:

-You're non-satisfied with your job and also not on high load.
-You're on high load and really pissed off.
-You just retire.

Well, that's it for now. :)