---
title: 'some nets!'
date: "`r Sys.Date()`"
output:
  html_document:
    fig_height: 6
    fig_width: 8
    highlight: pygments
    theme: readable
    toc: yes
---
#Intro
OK, we did look at data, now load this shit up and will try to make NNet.


Hello there! In this notebook I will look at popular HR dataset and will try to solve some problems and draw some graphs.

#Prerequisites

Libraries
```{r libs, message=FALSE, warning=FALSE}
library(ggplot2)
library(dplyr)
library(magrittr)
library(reshape2)
library(broom)
library(mlr)
library(keras)
```

Now load data, fix variables a bit:
```{r data, message=FALSE, warning=FALSE}
hr_data <- read.csv('../data/HR_comma_sep.csv')
hr_data %<>% 
  rename(department = sales, 
         work_accident = Work_accident, 
         average_monthly_hours = average_montly_hours,
         time_spent_company = time_spend_company) %>%
  mutate(promotion_last_5years = factor(promotion_last_5years, levels = c(1, 0), labels = c('Yes', 'No')),
         left = factor(left, levels = c(1, 0), labels = c('Yes', 'No')),
         work_accident = factor(work_accident, levels = c(1, 0), labels = c('Yes', 'No')))

```

#keRas

So, now that we have Keras in R, we can do some serious shit here.
This one for model structure
```{r}
net <- keras_model_sequential()
net %<>% 
  layer_dense(units = 10, activation = 'relu', input_shape = c(9)) %>%
  layer_dropout(rate = .5) %>%
  layer_dense(units = 5, activation = 'relu') %>%
  layer_dropout(rate = .2) %>%
  layer_dense(units = 1, activation = 'sigmoid')
```

Now compile model.
```{r compile}
net %<>% compile(loss = 'binary_crossentropy',
                 optimizer = optimizer_adam(lr = .001),
                 metrics = c('accuracy'))
```

And now we'll make data, and run model. Yay~

```{r data training}
sample <- sample(nrow(hr_data), .8*nrow(hr_data))
d.train <- hr_data %>% 
             mutate(work_accident = as.numeric(work_accident),
                    left = as.integer(left),
                    promotion_last_5years = as.numeric(promotion_last_5years),
                    department = as.numeric(department),
                    salary = as.numeric(salary)) %>% slice(sample)
d.test <- hr_data %>% 
             mutate(work_accident = as.numeric(work_accident),
                    left = as.integer(left),
                    promotion_last_5years = as.numeric(promotion_last_5years),
                    department = as.numeric(department),
                    salary = as.numeric(salary)) %>% slice(-sample)
train.x <- d.train %>% select(-left)
train.x %<>% t %>% matrix(ncol = 9)
train.y <- matrix(t(d.train$left)) - 1
```

```{r fit, message=FALSE, warning=FALSE}
history <- net %>% fit(train.x, train.y, epochs = 10, validation_split = .3)
```

```{r}
plot(history) + theme_bw()
```

Well, not really gud, but could be worse and also, it's our 1st net! Yay!